---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# bowl18: Implementation of Bowling et al.'s (2018) Vocal Similarity Algorithms

[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis build status](https://travis-ci.org/pmcharrison/bowl18.svg?branch=master)](https://travis-ci.org/pmcharrison/bowl18)
[![Coverage status](https://coveralls.io/repos/github/pmcharrison/bowl18/badge.svg)](https://coveralls.io/r/pmcharrison/bowl18?branch=master)

The `bowl18` R package implements Bowling et al.'s (2018)
vocal similarity algorithms.

## Installation

Within R, you can install the current version of `bowl18` from Github as follows:

``` r
if (!require(devtools)) install.packages("devtools")
devtools::install_github("bowl18")
```

## Usage

Bowling et al. (2018) formalise two aspects of a musical chord's
'vocal similarity':

a) Its similarity to the harmonic series, 
assessed with the `gill09_harmonicity()` function;
b) The absence of small frequency intervals,
assessed with the `bowl18_min_freq_dist()` function.

### gill09_harmonicity

`gill09_harmonicity()` measures a chord's similarity to a harmonic series,
after Gill & Purves (2009).
It assumes that chord pitches are precisely 
aligned with the just-tuned scale provided by Bowling et al. (2018).

By default, `gill09_harmonicity` assumes that the input 
is a vector of MIDI pitch numbers.
We recommend using the `hrep` package to clarify this input format, as follows:

```{r}
library(bowl18)
library(hrep)

# Constructing chords with pi_chord()
pi_chord(c(60, 64, 67)) # C major chord

# Major chord
gill09_harmonicity(pi_chord(c(60, 64, 67)))

# Diminished chord
gill09_harmonicity(pi_chord(c(60, 63, 66)))

# Augmented chord
gill09_harmonicity(pi_chord(c(60, 64, 68)))
```

### bowl18_min_freq_dist

`bowl18_min_freq_dist()` returns the minimum frequency distance between
the fundamental frequencies of a chord.
It makes no assumptions about the chord's tuning.

By default, `bowl18_min_freq_dist()` assumes that the input 
is a vector of frequencies.
We recommend using the `hrep` package to clarify this input format, as follows:

```{r}
library(bowl18)
library(hrep)

# Constructing chords with pi_chord() and fr_chord()
x <- pi_chord(c(60, 64, 67)) # C major chord
fr_chord(x) 

# Major chord
bowl18_min_freq_dist(fr_chord(pi_chord(c(60, 64, 67))))

# Cluster chord
bowl18_min_freq_dist(fr_chord(pi_chord(c(60, 61, 62))))
```

Bowling et al. (2018) argue that, in particular, frequency differences smaller
than 50 Hz contribute negatively to consonance.

## References

Bowling, D. L., Purves, D., & Gill, K. Z. (2018). Vocal similarity predicts the relative attraction of musical chords. Proceedings of the National Academy of Sciences, 115(1), 216–221. https://doi.org/10.1073/pnas.1713206115

Gill, K. Z., & Purves, D. (2009). A biological rationale for musical scales. PLoS ONE, 4(12), 1–9. https://doi.org/10.1371/journal.pone.0008144
